---
title: "Cryptography Course Advertisement Ad Analysis"
author: "Brian Onyango"
date: '2022-05-27'
output:
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Overview
A Kenyan entrepreneur has created an online cryptography course and would want to advertise it on her blog. She currently targets audiences originating from various countries. In the past, she ran ads to advertise a related course on the same blog and collected data in the process. She would now like to employ your services as a Data Science Consultant to help her identify which individuals are most likely to click on her ads. 

## Specifying the Question
Which individuals are most likely to clink on course advertisement ads?

## Defining the Metric for Success
This project will be successful if we will be able to determine factors that lead to a user to click an ad.

## Understanding the Context
Ad Clicks, or simply Clicks, is a marketing metric that counts the number of times users have clicked on a digital advertisement to reach an online property.If you have a campaign running, you are probably able to access click data on each specific ad. You may see data like this: Ad 1: 4,686 clicks Ad 2: 1,248 clicks Ad 3: 984 clicks You can see that Ad 1 is the higher performing ad by clicks. You may want to evaluate this ad and figure out why audiences tend to click on it more. You may also want to review Ad 3 and try to determine why it is not receiving as many clicks.

## Recording the Experimental Design

1. Data sourcing/loading
2. Data Understanding
3. Data Relevance
4. External Dataset Validation
5. Data Preparation
6. Univariate Analysis
7. Bivariate Analysis
8. Multivariate Analysis
9. Conclusion
10. Recommendations

## Data Relevance
The dataset avialed the client can be downloaded from this link <http://bit.ly/IPAdvertisingData>

Loading the Dataset
```{r}
df <- read.csv('http://bit.ly/IPAdvertisingData')
```
Previewing the top of our dataset
```{r}
head(df)
```
Previewing the bottom of the dataset
```{r}
tail(df)
```
Checking the data types
```{r}
# Data set structure.
str(df)
```
## Data Preparation
### Validity
```{r}
# checking for unnecesary columns 
colnames(df)
```
 columns seems necessary for this study
```{r}
# Checking for anomalies
summary(df)
```
 there are no anomalies in the data set
 
### Consistency
```{r}
# checking for missing values
colSums(is.na(df))
```
 there are no missing values
 
### Completeness
```{r}
# checking for duplicates
sum(duplicated(df))
```
there are no duplicates in this dataset

### Uniformity
```{r}
# Checking column names uniformity
colnames(df)
```
the column names are uniform

### Outliers
Using boxplots to check for outliersin numerical columns
```{r}
# Area.Income column
boxplot(df$Area.Income, data=df, main ="Area Income", ylab = 'Count', horizontal = TRUE)
```
```{r}
# Daily.Time.Spent.on.Site column
boxplot(df$Daily.Time.Spent.on.Site, data=df, main ="Daily.Time.Spent.on.Site", ylab = 'Count', horizontal = TRUE)
```
```{r}
# Age
boxplot(df$Age, data=df, main ="Age", ylab = 'Count', horizontal = TRUE)
```
```{r}
# Daily.Internet.Usage
boxplot(df$Daily.Internet.Usage, data=df, main ="Daily.Internet.Usage", ylab = 'Count', horizontal = TRUE)
```
from the boxplots, Area income column has outliers but we will keep them for futher analysis
## Exploratory Data Analysis
### Univariate Analysis
#### Categorical Analysis

Analysis using Countplots
```{r}
# 
library(ggplot2)
ggplot(df, aes(x=Male)) + geom_bar(fill=rgb(0.4,0.1,0.5))
```
ther are more females than males 
```{r}
ggplot(df, aes(x=factor(`Clicked.on.Ad`))) + geom_bar( fill=rgb(0.6,0.4,0.4))
```
the number of people who clicked the ad is equal to those who did not click

Analysis using Histograms
```{r}
# Age column
hist(df$Age)
```
```{r}
# Area.Income column
hist(df$Area.Income)
```
```{r}
# Daily.Time.Spent.on.Site column
hist(df$Daily.Time.Spent.on.Site)
```
```{r}
# Daily.Internet.Usage column
hist(df$Daily.Internet.Usage)
```
#### Numerical Analysis
Measures of Central Tendency
Mean
```{r}
# Mean of all numeric columns
colMeans(df[sapply(df,is.numeric)])
```
Median
```{r}
# Median of Daily.Time.Spent
median <- median(df$Daily.Time.Spent.on.Site)
print(median)
```
```{r}
# Median of Age
median <- median(df$Age)
print(median)
```
```{r}
# Median of Area.Income
median <- median(df$Area.Income)
print(median)
```
```{r}
# Median of Area.Income
median <- median(df$Daily.Internet.Usage)
print(median)
```
Mode
```{r}
# Creating the mode function
getmode <- function(v) {
   uniqv <- unique(v)
   uniqv[which.max(tabulate(match(v, uniqv)))]}
```

```{r}
# Age Mode
getmode(df$Age)
```
```{r}
# Daily.Time.Spent.on.Site Mode
getmode(df$Daily.Time.Spent.on.Site)
```
```{r}
# Area.Income Mode
getmode(df$Area.Income)
```
```{r}
# Daily.Internet.Usage Mode
getmode(df$Daily.Internet.Usage)
```
```{r}
# City Mode
getmode(df$City)
```
```{r}
# Ad.Topic.Line Mode
getmode(df$Ad.Topic.Line)
```
```{r}
# Country Mode
getmode(df$Country)
```
```{r}
# Timestamp Mode
getmode(df$Timestamp)
```
Varience
```{r}
# varience in Age
var(df$Age)
```
```{r}
# Daily.Time.Spent.on.Site varience
var(df$Daily.Time.Spent.on.Site)
```
```{r}
# Area.Income varience
var(df$Area.Income)
```
```{r}
# Daily.Internet.Usage varience
var(df$Daily.Internet.Usage)
```
Standard Deviation
```{r}
# Age SD
sd(df$Age)
```
```{r}
# Daily.Time.Spent.on.Site SD
sd(df$Daily.Time.Spent.on.Site)
```
```{r}
# Area.Income SD
sd(df$Area.Income)
```
```{r}
# Daily.Internet.Usage SD
sd(df$Daily.Internet.Usage)
```
Quantiles
```{r}
# Age quantiles
quantile(df$Age)
```
```{r}
# Daily.Time.Spent.on.Site quantiles
quantile(df$Daily.Time.Spent.on.Site)
```
```{r}
# Area.Income quantiles
quantile(df$Area.Income)
```
```{r}
# Daily.Internet.Usage quantiles
quantile(df$Daily.Internet.Usage)
```
### Bivariate Analysis

#### Pair Ggplots
```{r}
# Clicked.on.Ad vs Area.Income
ggplot(data = df, aes(x = Area.Income, fill = Clicked.on.Ad))+
        geom_histogram(bins  =20,col = "orange")+
        labs(title = "Income Distribution", x = "Area Income", y= "Frequency", fill = "Clicked on Ad")+ scale_color_brewer(
                palette = "Set1"
        )
```
```{r}
# Clicked.on.Ad vs Age
ggplot(data = df, aes(x = Age, fill = Clicked.on.Ad))+
        geom_histogram(bins  =20,col = "orange")+
        labs(title = "Age Distribution", x = "Age", y= "Frequency", fill = "Clicked on Ad")+ scale_color_brewer(
                palette = "Set1"
        )
```
```{r}
# Clicked.on.Ad vs Daily.Time.Spent.on.Site
ggplot(data = df, aes(x =Daily.Time.Spent.on.Site, fill = Clicked.on.Ad))+
        geom_histogram(bins  =20,col = "orange")+
        labs(title = "Daily Time Spent on Site", x = "Time Spent on Site", y= "Frequency", fill = "Clicked on Ad")+ scale_color_brewer(
                palette = "Set1"
        )
```
#### Scatter Plots
```{r}
# Scatter plot and correlation function
scatter.plt <- function(col1, col2, corr1, corr2, data, title){
data <- ggplot(data, aes(x = {{col1}}, y= {{col2}})) + geom_point(color = '#281E5D') + ggtitle(paste(title, 'Scatter Plot')) + theme(plot.title = element_text(hjust = 0.5))
correlation <- cor(df[ , c(corr1)], df[, c(corr2)])
plot(data)
print(paste0('Correlation = ', correlation, '.'))
}
```

Age Vs Daily Internet Usage
```{r}
scatter.plt(Daily.Internet.Usage, Age, data = df, corr1 = 'Daily.Internet.Usage', corr2 = 'Age', 'Age Vs Daily Internet Usage')
```
Age Vs Daily Time Spent on Site
```{r}
scatter.plt(Daily.Time.Spent.on.Site, Age, data = df, corr1 = 'Daily.Time.Spent.on.Site', corr2 = 'Age', 'Age Vs Daily Time Spent on Site')
```
Age Vs Area Income
```{r}
scatter.plt(Area.Income, Age, data = df, corr1 = 'Area.Income', corr2 = 'Age', 'Age Vs Area Income')
```
### Multivariate Analysis

#### Correlation matrix


```{r}
library(ggcorrplot)
corr <- dplyr::select(df,Age,Area.Income,Clicked.on.Ad,Daily.Internet.Usage,Daily.Time.Spent.on.Site,Male )
ggcorrplot(cor(corr), lab = TRUE, title = 'Correlation Heatmap', colors = c('#022D36', 'white', '#48AAAD'))
```

#### Scatter Plots
```{r}
ggplot(df, aes(Area.Income,Age))+geom_point(aes(colour= factor(`Clicked.on.Ad`)))+
  labs(title = "Scatter Plot of Age Distribution vs Area Income",
       x = "Area Income",
       y = "Age")
```

```{r}
ggplot(df, aes(Area.Income, Daily.Internet.Usage))+
  geom_point(aes(colour= factor(`Clicked.on.Ad`)))+
  labs(title = "Scatter Plot of Area Income vs Daily Internet Usage",
       x = "Area Income",
       y = "Daily Internet Usage")
```
```{r}
ggplot(df, aes(Age, Daily.Time.Spent.on.Site))+
  geom_point(aes(colour= factor(`Clicked.on.Ad`)))+
  labs(title = "Scatter Plot of Age Distribution vs Time Spent on Site",
       x = "Age",
       y = "Time Spent on Site")
```

```{r}
ggplot(df, aes(Daily.Time.Spent.on.Site, Area.Income))+
  geom_point(aes(colour= factor(`Clicked.on.Ad`)))+
  labs(title = "Time spent on site vs Income",
       x = "Daily Time Spent on Site",
       y = "Income Distribution")
```

```{r}
ggplot(df, aes(Age, Daily.Internet.Usage))+
  geom_point(aes(colour= factor(`Clicked.on.Ad`)))+
  labs(title = "Scatter Plot of Age Distribution vs Daily Usage",
       x = "Age",
       y = "Daily Usage")
```

### Conclusion
In conclusion, from the analysis, the major factors that determine if a user will click an ad are the:

1. Gender
2. Daily time spent on the site
3. Area Income
4. Time of day and month

From the analysis, we can conclude that:

1. The Older people (above 35),were more likely to click on the course advert.
2. The higher the person earns the less likely he/she will click the add.
3. There is an equal chance for someone to either click on the advert or not
4. The amount of time someone spent on the blog was inversely proportional to the probability of him/her clicking the add.

### Recommendations
From the analysis, I will recommend that:

1. More adverts to be targeted to older people above 35
2. The adverts to target low income (less than 60,000) population.
3. The adverts to be designed in such a way that a user clearly sees it when he/she access the blog
